---
title: "Notes 9: Application (Wikiart)"
output: html_document
---

```{r, message=FALSE}
library(readr)
library(dplyr)
library(ggplot2)
library(stringi)
library(jpeg)
library(cleanNLP)
library(igraph)

theme_set(theme_minimal())
source("scripts/funs.R")
```

# Data and EDA

As a final case-study, we can look at a collection of paintings from WikiArt:

```{r, message=FALSE}
wiki <- read_csv(file.path("data", "wikiart.csv"))
wiki
```

We have included paintings from three people. Each is also tagged with a "genre" code, which we can
explore here:

```{r, message=FALSE}
ggplot(wiki, aes(genre)) +
  geom_bar(aes(fill = artist)) +
  coord_flip() +
  scale_fill_viridis_d()
```

Do the results here match your prior understanding (if any) of these artists?

# Color analysis

Using color analysis seems particularly appropriate when looking at paintings. Lets see
some of the most blue:

```{r}
show_image(file.path("images", "wikiart", top_n(wiki, avg_blue, n=15)$path), ncol=5)
```


```{r, message=FALSE}
ggplot(wiki, aes(artist, avg_blue)) +
  stat_summary(fun.data=mean_cl_boot)
```

And most red images:

```{r}
show_image(file.path("images", "wikiart", top_n(wiki, avg_red, n=15)$path), ncol=5)
```

Also, feel free to try some other colors as well as the value and saturation. We can even
look at an analysis of the colors by artist, as we did with the posters:

```{r, message=FALSE}
ggplot(wiki, aes(artist, avg_red)) +
  stat_summary(fun.data=mean_cl_boot)
```

Try a different color and see what happens!

```{r}
show_image(file.path("images", "wikiart", top_n(wiki, avg_green, n=15)$path), ncol=5)
```



```{r, message=FALSE}
ggplot(wiki, aes(artist, avg_green)) +
  stat_summary(fun.data=mean_cl_boot)
```

```{r}
show_image(file.path("images", "wikiart", top_n(wiki, avg_yellow, n=15)$path), ncol=5)
```

```{r, message=FALSE}
ggplot(wiki, aes(artist, avg_yellow)) +
  stat_summary(fun.data=mean_cl_boot)
```
```{r}
show_image(file.path("images", "wikiart", top_n(wiki, avg_violet, n=15)$path), ncol=5)
```


```{r, message=FALSE}
ggplot(wiki, aes(artist, avg_violet)) +
  stat_summary(fun.data=mean_cl_boot)
```

# Neural networks

We will also use neural networks. However, keep in mind that these are all trained on photographs.
It is not instantly clear that they can be directly applied to paintings.

```{r, message=FALSE}
resnet <- read_csv(file.path("models", "wikiart-resnet50.csv"))
resnet
```

What are some of the most common objects:

```{r}
filter(resnet, rank == 1) %>%
  count(class_description, sort=TRUE)
```

As with the other collections, the most common ones seem like mistakes, but there are
some objects that may actually be present in the dataset. For example, the "lakeside"
category:

```{r}
index <- which(resnet$rank == 1 & resnet$class_description == "lakeside")
show_image(file.path("images", "wikiart", resnet$path[index]), ncol=5)
```

And hoopskirts (more-or-less):

```{r}
index <- which(resnet$rank == 1 & resnet$class_description == "hoopskirt")
show_image(file.path("images", "wikiart", resnet$path[index]), ncol=3)
```

Not all of the objects work as expected, however. The category of "groom" gives
particularly bad and confusing results:

```{r}
index <- which(resnet$rank == 1 & resnet$class_description == "groom")
show_image(file.path("images", "wikiart", resnet$path[index]), ncol=5)
```

While the object detection does not work quite as well on paintings, the image embedding algorithm
performs quite well. 

```{r}
embed <- read_rds(file.path("models", "wikiart-pool.rds"))
dist_mat <- as.matrix(dist(embed))
nearest_img <- t(apply(dist_mat, 2, function(v) order(v)[seq(1, 15)]))
```

Start with, for example, image 2:

```{r}
start_img <- 411
show_image(file.path("images", "wikiart", wiki$path[nearest_img[start_img,]]), ncol=5)
```
```{r}
start_img <- 411
print(file.path("images", "wikiart", wiki$path[nearest_img[start_img,]]))
```


Try using other starting images and see what happens. Does it make sensible suggestions? What
aspects of the paintings are most commonly used to find similar images?
